The Effect of Different Writing Tasks on Linguistic Style:\\ A Case Study of the ROC Story Cloze Task

### 2017-02-06 docs/language_constraint.tex
Roy V4.5

We train a logistic regression classifier to categorize an ending,
either as right vs.~wrong or as original vs.~new.
Each feature vector is computed using the words in one ending, without considering earlier parts of the story. 
We use the following style features.

@@

        \legend{Random Baseline, Our Model}
    \end{axis}
\end{tikzpicture}
\caption{\label{results} Results of  experiments 1 and 2 (left two charts). 
Rightmost graph shows a control experiment which classifies {\it
  original} endings vs \textit{\textbf{wrong}}
endings. \nascomment{this is overkill.  I suggest replacing it with a
  table that just shows our model, no baseline, and reiterate in the
  caption that our setup implies a 50\%-accurate random baseline.
  Also, I like including the third comparison, but I don't
  see how it's a ``control.'' }}
\end{figure}
}

\begin{table}[!t]
\begin{center}
%\small
\begin{tabular}{|l|r|} \hline
{\bf Experiment} & {\bf Accuracy} \\ \hline
{\sl right vs. wrong} & 0.645 \\ \hline
{\sl original vs. right} & 0.685 \\ \hline
{\sl original vs. wrong} & 0.756 \\ \hline
\end{tabular}
\end{center}
\caption{\label{results} Results of  experiments 1 ({\it right
    vs.~wrong}) and 2 ({\it original vs.~right (new)}). 
The bottom row shows an additional experiment which classifies {\it
  original vs.~wrong (new)} endings.
In all cases, our setup implies a 50\% random baseline.}
\end{table}

@@

\begin{table}[!t]
\begin{center}
%\small
\begin{tabular}{|l|r|} \hline
{\bf Model} & {\bf Acc.} \\ \hline
$\dagger${DSSM} \cite{Mostafazadeh:2016} & 0.585 \\ 
$\dagger${LexVec} \cite{Salle:2016} & 0.599 \\ \hline\hline
%{Niko (shared task)}	& 0.700\\ 
%{tbmihaylov (shared task)} & 0.711\\ \hline\hline
$\dagger${RNN}		& 0.677 \\ 
{\bf Ours} & {\bf 0.724} \\ 
$\dagger${\bf Combined (ours + RNN)} & {\bf 0.752} \\ \hline\hline
$\dagger$Human judgment & 1.000 \\ \hline
\end{tabular}
\end{center}
\caption{\label{cloze_results}\resolved{\yc{we should change accuracy to be in 0-100\% scale to be consistent with Table 2}}
Results on the test set of the  story cloze task. 
The first block are published results, 
the second block is our results.
LexVec results are taken from \cite{Speer:2016}.
%RNN is our implementation. 
Human judgement scores are taken from \cite{Mostafazadeh:2016}.
Methods marked with ($\dagger$) use the story context in order to make a prediction. 
 \resolved{\nascomment{I added 0s to ``Niko'' result
  and ``human''
  so sig. digits would line up.  if we don't have sig digits for some
  of them, make the zeros white so the numbers line up.  where did the
human judgments come from?  those are not mentioned in our text anywhere!}
}}
\end{table}

@@

Finally, when comparing the two groups of salient features from both experiments, we find an interesting trend.
Several features, such as ``\textsc{start} NNP'' and  ``NN .'', which indicate {\it wrong} sentences in Experiment 1, are used to predict {\it new} (i.e., {\it correct}) endings in Experiment 2. 
This indicates that, for instance, incoherent endings have a stronger tendency to begin with a proper noun compared to coherent endings, 
which in turn are more likely to do so than original endings. 
This partially explains why distinguishing between {\it original} and {\it wrong} endings is an easier task compared to the other pairs (\secref{Results}).

@@

\end{tabular}
\end{center}
\caption{\label{exp2_features}
The top 5 most heavily weighted features for predicting {\it original} vs.~{\it new} ({\it right}) endings.
}
\end{table}

@@

\paragraph{The effect of writing tasks on mental states.}
In this paper we have shown that giving a writer different writing tasks affects writing style in easily detected ways.
Our results indicate that when authors are asked to write the last
sentence of a five-sentence story, they will use different style to
write a {\it right} ending compared to a {\it wrong} ending. We have
also shown that writing the ending as part of one's own five-sentence story is very different than reading four sentences and then writing the fifth.
Our findings hint that the nature of the writing task imposes a
different mental state on the author, which is expressed in ways that can be observed using extremely simple automatic tools.

@@

\paragraph{The story cloze task.}
Our study also provides important insights for the future design of NLP tasks. 
The story cloze task was very carefully designed. Many factors, such
as topic diversity and temporal and causal relation diversity,  were controlled for \cite{Mostafazadeh:2016}. 
The authors also made sure each pair of endings was written by the
same author, partly in order to avoid author-specific style effects.
 Nonetheless, despite these efforts, several significant style
 differences can be found between the training and the test set, as well as between the positive and negative labels.

@@

\paragraph{Machine reading.}
The  story cloze task, which is the focus of this paper, is part of a wide set of machine reading/comprehension challenges published in the last few years.
These include datasets like bAbI \cite{Weston:2015}, SNLI \cite{bowman2015large}, CNN/DailyMail \cite{hermann2015teaching}, SQuAD \cite{rajpurkar2016squad}, and LAMBADA \cite{Paperno:2016}. 
While these works have presented resources for researchers, 
it is often the case that these datasets suffer from methodological
problems caused by applying noisy automatic tools to generate them \cite{Chen:2016}. 
In this paper, we have pointed to another methodological challenge in
designing machine reading tasks:  different writing tasks
used to generated the data affect writing style, confounding classification problems.
%, as well as create a discrepancy between the training and the test data.
 \resolved{\nascomment{dropped the train/test discrepancy here}}

>>>>

We train a logistic regression classifier to categorize an ending,
either as {\it right} vs.~{\it wrong} or as {\it original} vs.~new ({\it right}).
Each feature vector is computed using the words in one ending, without considering earlier parts of the story. 
We use the following style features.

@@

        \legend{Random Baseline, Our Model}
    \end{axis}
\end{tikzpicture}
\caption{\label{results}Results of  experiments 1 and 2 (left two charts). 
Rightmost graph shows a control experiment which classifies {\it
  original} endings vs \textit{\textbf{wrong}}
endings. \nascomment{this is overkill.  I suggest replacing it with a
  table that just shows our model, no baseline, and reiterate in the
  caption that our setup implies a 50\%-accurate random baseline.
  Also, I like including the third comparison, but I don't
  see how it's a ``control.'' }}
\end{figure}
}

\begin{table}[!t]
\begin{center}
%\small
\begin{tabular}{|l|r|} \hline
{\bf Experiment} & {\bf Accuracy} \\ \hline
{\sl right vs. wrong} & 0.645 \\ \hline
{\sl original vs. right} & 0.685 \\ \hline
{\sl original vs. wrong} & 0.756 \\ \hline
\end{tabular}
\end{center}
\caption{\label{results}Results of  experiments 1 ({\it right
    vs.~wrong}) and 2 ({\it original vs.~right (new)}). 
The bottom row shows an additional experiment which classifies {\it
  original vs.~wrong (new)} endings.
In all cases, our setup implies a 50\% random baseline.}
\end{table}

@@

\begin{table}[!t]
\begin{center}
%\small
\begin{tabular}{|l|r|} \hline
{\bf Model} & {\bf Acc.} \\ \hline
$\dagger${DSSM} \cite{Mostafazadeh:2016} & 0.585 \\ 
$\dagger${LexVec} \cite{Salle:2016} & 0.599 \\ \hline\hline
%{Niko (shared task)}	& 0.700\\ 
%{tbmihaylov (shared task)} & 0.711\\ \hline\hline
$\dagger${RNN}		& 0.677 \\ 
{\bf Ours} & {\bf 0.724} \\ 
$\dagger${\bf Combined (ours + RNN)} & {\bf 0.752} \\ \hline\hline
$\dagger$Human judgment & 1.000 \\ \hline
\end{tabular}
\end{center}
\caption{\label{cloze_results}\resolved{\yc{we should change accuracy to be in 0-100\% scale to be consistent with Table 2}}Results on the test set of the  story cloze task. 
The first block are published results, 
the second block is our results.
LexVec results are taken from \cite{Speer:2016}.
%RNN is our implementation. 
Human judgement scores are taken from \cite{Mostafazadeh:2016}.
Methods marked with ($\dagger$) use the story context in order to make a prediction. 
 \resolved{\nascomment{I added 0s to ``Niko'' result
  and ``human''
  so sig. digits would line up.  if we don't have sig digits for some
  of them, make the zeros white so the numbers line up.  where did the
human judgments come from?  those are not mentioned in our text anywhere!}
}}
\end{table}

@@

Finally, when comparing the two groups of salient features from both experiments, we find an interesting trend.
Several features, such as ``\textsc{start} NNP'' and  ``NN .'', which indicate {\it wrong} sentences in Experiment 1, are used to predict new (i.e., {\it right}) endings in Experiment 2. 
This indicates that, for instance, incoherent endings have a stronger tendency to begin with a proper noun compared to coherent endings, 
which in turn are more likely to do so than original endings. 
This partially explains why distinguishing between {\it original} and {\it wrong} endings is an easier task compared to the other pairs (\secref{Results}).

@@

\end{tabular}
\end{center}
\caption{\label{exp2_features}
The top 5 most heavily weighted features for predicting {\it original} vs.~new ({\it right}) endings.
{\it length} is the sentence length feature (see \secref{Model}).
}
\end{table}

@@

\paragraph{The effect of writing tasks on mental states.}
In this paper we have shown that giving a writer different writing tasks affects her writing style in easily detected ways.
Our results indicate that when authors are asked to write the last
sentence of a five-sentence story, they will use different style to
write a {\it right} ending compared to a {\it wrong} ending. We have
also shown that writing the ending as part of one's own five-sentence story is very different than reading four sentences and then writing the fifth.
Our findings hint that the nature of the writing task imposes a
different mental state on the author, which is expressed in ways that can be observed using extremely simple automatic tools.

@@

\paragraph{Design of NLP tasks.}
Our study also provides important insights for the future design of NLP tasks. 
The story cloze task was very carefully designed. Many factors, such
as topic diversity and temporal and causal relation diversity,  were controlled for \cite{Mostafazadeh:2016}. 
The authors also made sure each pair of endings was written by the
same author, partly in order to avoid author-specific style effects.
 Nonetheless, despite these efforts, several significant style
 differences can be found between the training and the test set, as well as between the positive and negative labels.

@@

\paragraph{Machine reading.}
The  story cloze task, which is the focus of this paper, is part of a wide set of machine reading/comprehension challenges published in the last few years.
These include datasets like bAbI \cite{Weston:2015}, SNLI \cite{bowman2015large}, CNN/DailyMail \cite{hermann2015teaching}, LAMBADA \cite{Paperno:2016} and SQuAD \cite{rajpurkar2016squad}. 
While these works have presented resources for researchers, 
it is often the case that these datasets suffer from methodological
problems caused by applying noisy automatic tools to generate them \cite{Chen:2016}. 
In this paper, we have pointed to another methodological challenge in
designing machine reading tasks:  different writing tasks
used to generated the data affect writing style, confounding classification problems.
%, as well as create a discrepancy between the training and the test data.
 \resolved{\nascomment{dropped the train/test discrepancy here}}
